---
title: "simpleRegress"
author: "sierra lander"
date: "Summer 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```
for model length~neck, adjusted r^2 = 0.75

## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 


## length based on age 



```{r}
age <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(age)
summary.lm(age)
```
for model length~age r^2 = .51, this is less than the adjusted r^2 for neck so this model is not as good as the neck model  

