---
title: "simpleRegress"
author: "sierra lander"
date: "Summer 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```
for model length~neck, adjusted r^2 = 0.75

## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 


## length based on age 



```{r}
age <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(age)
summary.lm(age)
```
for model length~age r^2 = .51, this is less than the adjusted r^2 for neck so this model is not as good as the neck model  

## length based on head length 


```{r}
headL <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(headL)
summary.lm(headL)
```

Adjusted r^2 = .84. this is better than the r^2 of length~neck, this means that model length~headlen is better 


## length based on head width


```{r}
headw <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(headw)
summary.lm(headw)
```


adjusted r^2 = .53

the adjusted r^2 for headlen is better than headwth 
Length~HEADLEN is better than length~HEADWTH


## length based on chest 



```{r}
chest <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chest)
summary.lm(chest)
```

adjusted r^2 = 0.79 

The adjusted r^2 for head length is 0.84 
This shows that model length~headlen is better than model length~chest 





## length based on sex




```{r}
sex <- lm(LENGTH~SEX,data=bears)
plot(LENGTH~SEX,data=bears)
summary.lm(sex)
```


The adjusted r^2 = -0.0084
adjusted r^2 for length~headlen is .84 

this shows that length~headlen is better than length~sex



## length based on weight 

```{r}
weight <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(weight)
summary.lm(weight)
```


The adjusted r^2 = 0.74   
the adjusted r^2 for length~headlen is 0.84   
this shows that length~headlen is better than length~weight  



### Conclusion

In conclusion the best model is LENGTH~HEADLEN is the best model it had the highest Adjusted r^2 which was 0.84 and all other models had lower scores. 


